---
title: "R Notebook"
output: html_notebook
---

# Part 3: Drawing Conclusions from Data

## Z-testing in R (population variance is known)
IQ test:  
H0: mean IQ <= 100  
H1: mean IQ >  100  

```{r}
IQ_data <- c(100, 101, 104, 109, 125, 116, 105, 108, 110)
mean_IQdata <- mean(IQ_data)
# IQ is known to have a SD of 15
popsd <- 15
# test if mean of sample is statistically different from the 
# known IQ-mean of 100
mu <- 100
# Assume that alpha = 0.05
?pnorm
```

Because you use the z-score in the hypothesis test, the z-score here is called the **test statistic**. 

```{r Z score as test statistic}
z.score <- round((mean(IQ_data) - mu)/
                     (popsd / sqrt(length(IQ_data))), 3)
z.score
```


```{r ztest function}
z.test <- function(x, mu, popsd) {
    one.tail <- NULL
    z.score <- round((mean(x) - mu)/(popsd/sqrt(length(x))), 3)
    one.tail.p <- round(pnorm(abs(z.score), lower.tail = F), 3)
    cat(" z =", z.score, "\n",
        "one-tailed probability =", one.tail.p, "\n",
        "two-tailed probability =", 2*one.tail.p)
}

z.test(IQ_data, mu, popsd)
```

You set up a critical value, a decision criterion. The **critical value** is determined by alpha. 

```{r critical value as z-score}
qnorm(
    p = 0.95,
    lower.tail = T
)
```

```{r critical value as IQ points}
qnorm(
    p = 0.95,
    mean = 100,
    sd = popsd/sqrt(length(IQ_data)),
    lower.tail = T
)
```

With pnorm() only:

```{r}
pnorm(q = z.score,
      lower.tail = FALSE)
```

P-value is very low, this also suggests that H0 can be rejected.

Same using means and standard error instead of z-score:

```{r}
pnorm(
    q = mean_IQdata,
    mean = 100,
    sd = popsd/sqrt(length(IQ_data)),
    lower.tail = FALSE
)
```

Show graphically using tigerstat:

```{r}
library(tigerstats)
?tigerstats::pnormGC

pnormGC(
    bound = z.score,
    region = "above",
    mean = 0,
    sd = 1,
    graph = T
)
```

Tigerstat again, but using means instead of z-scores:

```{r}
pnormGC(
    bound = mean_IQdata,
    region = "above",
    mean = 100,
    sd = 15/sqrt(length(IQ_data)),
    graph = T
)
```

The **critical value** - the value of z that cuts off 5 percent of the area in a standard normal distribution - is 1.645:

```{r}
qnorm(p = 0.95)
```

```{r}
qnormGC(area = 0.95, graph = TRUE)
```

In case of a two-tailed test the 5% is divided evenly between the left tail and the right tail.  

```{r}
qnorm(p = 0.975)
```

```{r}
qnormGC(area = 0.975, graph = TRUE)
```


## t for One

You usually have small samples.
Usually you do NOT know the population parameters.
Use 't' as a **test statistic**.

```{r}
FarKlempt_data <- c(3,6,9,9,4,10,6,4,12)
?t.test
```

T-test:  

```{r}
t.test(FarKlempt_data,
       mu = 4,                      # H0 hypothesis
       alternative = "greater")
```

t-value calculation against an assumed (H0) mean of 4

```{r}
mean_H0 <- 4

t_far_klemp_data <-
    (mean(FarKlempt_data) - mean_H0) / 
    (sd(FarKlempt_data) / 
         sqrt(length(FarKlempt_data)))
```

```{r}
?pt

pt(q = 2.8823,
   df = length(FarKlempt_data)-1,
   lower.tail = F)
```


